### 2023-11-27 docs/methods.qmd
Draft the initial report

# Methods

- Determine confidence interval in ACF: https://stats.stackexchange.com/questions/211628/how-is-the-confidence-interval-calculated-for-the-acf-function