### 2024-01-26 docs/index.qmd
Explore monthly seasonality

# Executive Summary {-}

::: {.content-visible when-profile="seasonality"}

Exploration on seasonality is done on monthly data with a first-order difference

:::