### 2024-02-05 docs/index.qmd
Add templates for executive summaries

# Executive Summary {-}

::: {.content-visible when-profile="descriptive"}

{{< include exec-summary/descriptive.qmd >}}

:::

::: {.content-visible when-profile="arima"}

{{< include exec-summary/arima.qmd >}}

:::

::: {.content-visible when-profile="seasonality"}

{{< include exec-summary/seasonality.qmd >}}

:::