### 2024-02-07 docs/results/spectral.qmd
Report the reconstructed time series, close #9 CU-86cuhtr9g[closed], update #6 CU-86cuhtqmg[in-progress]

---
knitr:
  opts_chunk:
    fig.width: 14
    fig.height: 8
---

\KOMAoptions{paper=landscape,pagesize}
\recalctypearea

```{r init}

pkgs <- c("magrittr", "targets")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)

options(digits = 2)

reportResid <- function(mod) {

  group <- attr(mod, "group")
  res   <- residuals(mod)

  normality <- shapiro.test(res) %>%
    broom::tidy()

  tbl   <- res %>%
    tseries::pp.test() %>%
    broom::tidy() %>%
    subset(select = c("statistic", "p.value", "method")) %>%
    set_colnames(c("stat", "p", "method")) %>%
    inset(c("stat2", "p2", "method2"), value = normality) %>%
    tibble::add_column("group" = group, .before = 1) %>%
    dplyr::mutate(
      "mean"   = mean(res),
      "median" = median(res),
      "sd"     = sd(res),
      "IQR"    = IQR(res)
    )

  return(tbl)
}

```

# Results

## Residuals analysis

### Number of medication claims

```{r}

tar_read(mod_ssa_n_claim) %>%
  lapply(reportResid) %>%
  {do.call(rbind, .)} %>%
  knitr::kable()

```

### Claim-to-patient ratio

```{r}

tar_read(mod_ssa_claim2patient) %>%
  lapply(reportResid) %>%
  {do.call(rbind, .)} %>%
  knitr::kable()

```

### Eigenvector centrality

```{r}

tar_read(mod_ssa_eigen) %>%
  lapply(reportResid) %>%
  {do.call(rbind, .)} %>%
  knitr::kable()

```

{{< pagebreak >}}

## Reconstructed plots

### Number of medication claim

```{r}

tar_read(plt_ssa_recon_n_claim)

```

### Medication claim to patient ratio

```{r}

tar_read(plt_ssa_recon_claim2patient)

```

### Eigenvector centrality

```{r}

tar_read(plt_ssa_recon_eigen)

```