### 2024-02-09 docs/exec-summary/seasonality.qmd

### Exploring seasonality in the dataset

Exploration on seasonality is done on monthly data with a first-order difference.