### 2024-02-09 docs/exec-summary/spectral.qmd

### Decomposition with singular spectrum analysis