### 2024-02-09 docs/index.qmd

# Executive Summary {-}

```{r init}

options(scipen = 999)

```

::: {.content-visible when-meta="in.report"}
{{< include exec-summary/report.qmd >}}
:::

::: {.content-visible when-profile="descriptive"}
{{< include exec-summary/descriptive.qmd >}}
:::

::: {.content-visible when-profile="arima"}
{{< include exec-summary/arima.qmd >}}
:::

::: {.content-visible when-profile="seasonality"}
{{< include exec-summary/seasonality.qmd >}}
:::

::: {.content-visible when-profile="spectral"}
{{< include exec-summary/spectral.qmd >}}
:::

::: {.content-visible when-meta="in.report"}

## Contents

- @sec-descriptive contains a descriptive explanation of the dataset
- @sec-seasonality reports exploration of seasonality on detrended data
- @sec-arima showcases results obtained from ARIMA models
- @sec-spectral outlines the decomposed time series using singular-spectrum analysis

:::