### 2024-02-09 docs/results/spectral.qmd

---
knitr:
  opts_chunk:
    fig.width: 18
    fig.height: 12
---

```{r init}

pkgs <- c("magrittr", "targets", "kableExtra")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)

options(digits = 2, scipen = 999)

prettyRound <- function(x, ...) {
  x %<>% round(...) %>% {ifelse(. == 0, "<0.001", as.character(.))}
}

reportResid <- function(mod) {

  group <- attr(mod, "group")
  res   <- residuals(mod)

  normality <- shapiro.test(res) %>%
    broom::tidy()

  tbl <- res %>%
    tseries::pp.test() %>%
    broom::tidy() %>%
    subset(select = c("statistic", "p.value", "method")) %>%
    set_colnames(c("stat", "p", "method")) %>%
    inset(c("stat2", "p2", "method2"), value = normality) %>%
    tibble::add_column("group" = group, .before = 1) %>%
    dplyr::mutate(
      "Unit Root" = sprintf("%s (p %s)", round(stat, 2), prettyRound(p, 3)),
      "Normality" = sprintf("%s (p %s)", round(stat2, 2), prettyRound(p2, 3)),
      "Mean"   = mean(res),
      "SD"     = sd(res),
      "Median" = median(res),
      "IQR"    = IQR(res)
    ) %>%
    subset(select = -c(stat:method2)) %>%
    dplyr::rename("ATCC Group" = group)

  return(tbl)
}

```

\normalpapersize

# Results on singular spectrum analysis {#sec-spectral}

## Trend evaluation

```{r tbl-trend}
#| tbl-cap: Trend of medication uses alongside its measure of relative importance

fnote <- "The statistics were obtained from Mann-Kendall trend test, where positive result signifies a positive slope"

tar_read(decom_ssa) %>%
  subset(.$component %in% c("Original", "Trend") & .$metric != "claim2patient") %>%
  tibble::tibble() %>%
  dplyr::group_by(metric, group, component) %>%
  dplyr::group_modify(~trend::mk.test(.x$value) %>% broom::tidy()) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    "Mann-Kendall" = sprintf(
      "%s (p %s)", prettyRound(statistic, 2), prettyRound(p.value, 3)
    )
  ) %>%
  subset(select = -c(statistic:alternative)) %>%
  tidyr::pivot_wider(names_from = metric, values_from = `Mann-Kendall`) %>%
  tidyr::pivot_wider(names_from = component, values_from = eigen:n_claim) %>%
  set_colnames(c("ATCC Group", rep(c("Original", "Trend"), 2))) %>%
  knitr::kable(longtable = TRUE, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "repeat_header") %>%
  kableExtra::add_header_above(c(" " = 1, "Eigenvector Centrality" = 2, "Medication Claims" = 2)) %>%
  kableExtra::add_footnote(fnote, notation = "none")

```

\uselandscape

## Residuals analysis

### Number of medication claims

```{r tbl-resid-n-claim}
#| tbl-cap: Residuals analysis for unit-root and normality test for the decomposition model of number of claims

tar_read(mod_ssa_n_claim) %>%
  lapply(reportResid) %>%
  {do.call(rbind, .)} %>%
  knitr::kable(longtable = TRUE, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "repeat_header", font_size = 8)

```

### Eigenvector centrality

```{r tbl-resid-eigen}
#| tbl-cap: Residuals analysis for unit-root and normality test for the decomposition model of relative importance

tar_read(mod_ssa_eigen) %>%
  lapply(reportResid) %>%
  {do.call(rbind, .)} %>%
  knitr::kable(longtable = TRUE, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = "repeat_header", font_size = 8)

```

{{< pagebreak >}}

## Reconstructed plots {#sec-recon}

### Number of medication claim

```{r}

tar_read(plt_ssa_recon_n_claim)

```

::: {.content-visible when-profile="spectral"}

### Claim-to-patient ratio

```{r}

tar_read(plt_ssa_recon_claim2patient)

```

:::

### Eigenvector centrality

```{r}

tar_read(plt_ssa_recon_eigen)

```