### 2024-02-14 docs/exec-summary/_report.qmd

```{r}

pkgs <- c("targets", "magrittr")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)
options(digits = 2, scipen = 999)

```

## Abstract

{{< include abstract/_abstract.qmd >}}

:::: {.content-visible when-meta="in.report"}

{{< pagebreak >}}

## Analytical procedures

{{< include exec-summary/_source-data.qmd >}}

{{< include exec-summary/_preprocessing.qmd >}}

{{< include exec-summary/_measure-eigen.qmd >}}

{{< include exec-summary/_descriptive.qmd >}}

{{< include exec-summary/_seasonality.qmd >}}

{{< include exec-summary/_arima.qmd >}}

{{< include exec-summary/_spectral.qmd >}}

## Highlighted findings

{{< include exec-summary/_result-highlights.qmd >}}

::: {.content-visible when-profile="summary"}

\uselandscape

## Excerpt on findings

### Pair plot

```{r}

tar_read(plt_pair)[[17]]

```

### Dot plots on weekly time-series

```{r}

tar_read(plt_dot_week)[[1]]

```

### Differenced time-series on a weekly interval

```{r}

tar_read(plt_dot_diff_week)[[1]]

```

### Exploring seasonality within the number of medication claims

```{r}

tar_read(plt_period_month)[[1]]
tar_read(plt_acf_diff_month)[[1]]
tar_read(plt_pacf_diff_month)[[1]]

```

### Exploring ARIMA model

```{r}

tar_read(plt_arima_forecast_n_claim)[[17]]

```

### Decomposed time-series with SSA

```{r}

tar_read(plt_ssa_recon_n_claim)[[17]]

```

\normalpapersize

:::

::::