### 2024-02-14 docs/index.qmd

---
knitr:
  opts_chunk:
    fig.width: 18
    fig.height: 12
---

```{r}

pkgs <- c("targets", "magrittr")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)
options(digits = 2, scipen = 999)

```

# Executive Summary {-}

::: {.content-visible when-meta="in.report"}
{{< include exec-summary/_report.qmd >}}
:::

::: {.content-visible when-profile="descriptive"}
{{< include exec-summary/_descriptive.qmd >}}
:::

::: {.content-visible when-profile="arima"}
{{< include exec-summary/_arima.qmd >}}
:::

::: {.content-visible when-profile="seasonality"}
{{< include exec-summary/_seasonality.qmd >}}
:::

::: {.content-visible when-profile="spectral"}
{{< include exec-summary/_spectral.qmd >}}
:::

:::: {.content-visible when-meta="in.report"}

## Contents

- @sec-descriptive contains a descriptive explanation of the dataset
- @sec-seasonality reports exploration of seasonality on detrended data
- @sec-arima showcases results obtained from ARIMA models
- @sec-spectral outlines the decomposed time series using singular-spectrum analysis

## References

::: {#refs}
:::

::::