### 2024-03-21 docs/abstract/_abstract-eupha.qmd
Add abstract for EUPHA

---
title: Eigenvector centrality signifies antidepressants as a central hub of the co-medication network
---

Background: Mental health disorders are prevalent comorbidities affecting nearly two billion people worldwide. People with chronic diseases are at a higher risk of having mental health issues. We evaluate the interrelatedness of mental health problems with other diagnoses, reflected as a relative importance score of psychopharmaca in a medication claim registry.

Methods: Exploratory time-series analysis was performed in a static cohort from IADB.nl with at least one prescription of psychopharmaca from 2018 to 2022. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification. A medication-claim network was constructed to capture dose-adjusted co-prescription. Eigenvector centrality (ce) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims (nc) and eigenvector centrality were extracted as weekly-aggregated time-series data.

Results: Antidepressants (ce: 0.11, nc: 28,993), antipsychotics (ce: 0.07, nc: 14,027), and anxiolytics (ce: 0.08, nc: 14,061) are psychopharmaca with high eigenvector centrality. These classes of medication are likely to be co-prescribed alongside other medications. Other groups with high eigenvector centrality were medications for alimentary and metabolism (ce: 0.15, nc: 53,602), blood (ce: 0.06, nc: 13,310), cardiovascular (ce: 0.14, nc: 44,802), analgesics (ce: 0.09, nc: 13,847), and respiratory (ce: 0.09, nc: 25,034). Only antidepressants had a positive time trend for both eigenvector centrality and the number of claims.

Conclusion: Eight medications have a high concurrence on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants are related to co-prescription with other medication classes. We recommend constructing a medication claim network to evaluate periodic co-prescription as a public health monitoring procedure.