### 2024-04-18 docs/abstract/_abstract-eupha.qmd
Draft abstracts for ISPOR and EUPHA

**Background:**
Mental health disorders are prevalent comorbidities affecting nearly two billion people worldwide. People with chronic diseases are at a higher risk of having mental health issues. We evaluate the interrelatedness of mental health problems with other diagnoses, reflected as a relative importance score of psychopharmaca in a medication claim registry.

**Methods:**
Exploratory time-series analysis was performed in a static cohort from IADB.nl with at least one prescription of psychopharmaca from 2018 to 2022. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification. A drug prescription network was constructed to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims ($n_c$) and eigenvector centrality were extracted as weekly-aggregated time-series data.

**Results:**
Antidepressants ($c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`), antipsychotics ($c_e$: `r N05A$ce`, $n_c$: `r N05A$nc`), and anxiolytics ($c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) are psychopharmaca with high eigenvector centrality. These classes of medication are likely to be co-prescribed alongside other medications. Further groups with high eigenvector centrality were medications for alimentary and metabolism ($c_e$: `r A0x$ce`, $n_c$: `r A0x$nc`), blood ($c_e$: `r B0x$ce`, $n_c$: `r B0x$nc`), cardiovascular ($c_e$: `r C0x$ce`, $n_c$: `r C0x$nc`), analgesics ($c_e$: `r N02$ce`, $n_c$: `r N02$nc`), and respiratory ($c_e$: `r R0x$ce`, $n_c$: `r R0x$nc`). Only antidepressants had a positive time trend for both eigenvector centrality and the number of claims.

**Conclusion:**
Eight medications have a high concurrence on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants are related to co-prescription with other medication classes. We recommend constructing a medication claim network to evaluate periodic co-prescription as a public health monitoring procedure.