### 2024-04-18 docs/abstract/_abstract.qmd

```{r}

pkgs <- c("targets", "magrittr")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)
options(digits = 2, scipen = 999)

getGroup <- function(tbl, groupname) {
  sub_tbl <- tbl %>% subset(.$group == groupname, select = -group)
  return(sub_tbl)
}

results <- tar_read(ts_clust) %>%
  tibble::tibble() %>%
  dplyr::group_by(group) %>%
  dplyr::summarize(
    ce = mean(eigen),
    nc = mean(n_claim) %>% round() %>% scales::comma(),
    lim_ce = 1.96 * sd(eigen) / sqrt(length(eigen)),
    ce_ci  = sprintf("%.02f [%.02f, %.02f]", ce, ce - lim_ce, ce + lim_ce)
  )

N06A <- results %>% getGroup("Antidepressants")
N05A <- results %>% getGroup("Antipsychotics")
N05B <- results %>% getGroup("Anxiolytics")
A0x  <- results %>% getGroup("Alimentary and metabolism")
B0x  <- results %>% getGroup("Blood")
C0x  <- results %>% getGroup("Cardiovascular")
N02  <- results %>% getGroup("Analgesics")
R0x  <- results %>% getGroup("Respiratory")

```

::: {.content-visible when-meta="in.abstract.full"}

{{< include _abstract-full.qmd >}}

:::

::: {.content-visible when-meta="in.abstract.eupha"}

{{< include _abstract-eupha.qmd >}}

:::

::: {.content-visible when-meta="in.abstract.ispor"}

{{< include _abstract-ispor.qmd >}}

:::