### 2024-07-16 docs/abstract/_abstract-eupha.qmd
Initial draft for ISPOR 2024

**Background:**
Mental health disorders are prevalent comorbidities affecting nearly two billion people worldwide. People with chronic diseases are at a higher risk of having mental health issues. We evaluate the interrelatedness of mental health problems with other diagnoses, reflected as a relative importance score of psychopharmaca in a medication claim registry.

**Methods:**
Exploratory time-series analysis was performed in a closed cohort from IADB.nl with at least one prescription of anxiolytics/antidepressants from 2018 to 2022. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification. A drug prescription network was constructed to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims ($n_c$) and eigenvector centrality were extracted as weekly-aggregated time-series data.

**Results:**
Antidepressants ($c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics ($c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) are psychopharmaca with high eigenvector centrality. These classes of medication are likely to be co-prescribed alongside other medications. Further groups with high eigenvector centrality were medications for alimentary and metabolism ($c_e$: `r A0x$ce`, $n_c$: `r A0x$nc`), blood ($c_e$: `r B0x$ce`, $n_c$: `r B0x$nc`), cardiovascular ($c_e$: `r C0x$ce`, $n_c$: `r C0x$nc`), analgesics ($c_e$: `r N02$ce`, $n_c$: `r N02$nc`), and respiratory ($c_e$: `r R0x$ce`, $n_c$: `r R0x$nc`). Only antidepressants had a positive time trend for both eigenvector centrality and the number of claims.

**Conclusion:**
Seven medications, three of which are psychopharmaca, had a high eigenvector centrality, indicating concurrence with other medication groups on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants was related to co-prescription with other medication classes. Constructing a medication claim network enriches the exploratory analysis of medication use at the population level.