### 2024-08-14 docs/abstract/_abstract-full.qmd
Incorporate revision suggested by Spyros

**Background:**
Networks have been used in different scientific fields to model complex systems. In medicine and epidemiology, drug prescription networks (DPN) have been proposed as a modeling tool for studying the temporal dynamics of medication distribution and consumption within a population.

**Objectives:**  
Evaluate the structural properties of temporal DPNs that are constructed from the observed daily co-prescriptions between treatments for mental health problems and other drug therapies in a pharmacy dispensing database.

**Methods:**
For  exploratory time series analysis, and network construction, we utilized the IADB.nl database that collects prescriptions from 128 pharmacies in the Netherlands. A closed cohort with at least one prescription of anxiolytics/antidepressants dispensed anywhere from 2018 to 2022 was extracted. The drug prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification, where psychopharmaca was evaluated at ATC level 4, drugs for the nervous system at ATC level 3, and other medications at ATC level 1, distinguished into a total of 24 therapeutic medication classes. Symmetric daily dose-adjusted co-prescriptions were used to construct time-varying DPNs as undirected graphs. Subsequently, eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of prescriptions ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality were extracted as weekly-aggregated time-series data. Singular spectrum analysis was performed to decompose the data into its constituting trend, harmonics, and noise components.

**Results:**
Antidepressants (ATC `N06A`, $c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics (`N05B`, $c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) are psychopharmaca with high eigenvector centrality. These classes of medication are likely to be co-prescribed apart from other medications. Further groups with high eigenvector centrality were medications for alimentary and metabolism (`A01-A16`, $c_e$: `r A0x$ce`, $n_c$: `r A0x$nc`), blood (`B01-B06`, $c_e$: `r B0x$ce`, $n_c$: `r B0x$nc`), cardiovascular (`C01-C10`, $c_e$: `r C0x$ce`, $n_c$: `r C0x$nc`), analgesics (`N02`, $c_e$: `r N02$ce`, $n_c$: `r N02$nc`), and respiratory (`R01-R07`, $c_e$: `r R0x$ce`, $n_c$: `r R0x$nc`). Among the seven medication groups with high eigenvector centrality, only antidepressants showed a positive time trend for both eigenvector centrality and actual number of claims.

**Discussion:**
Seven medication therapy classes, three of which are psychopharmaca, had a high eigenvector centrality, indicating concurrence with other medication groups on a population level in the group of antidepressant or anxiolytic users. Positive time trends in both eigenvector centrality and dispensing suggest that increasing uses of antidepressants was related to co-prescription with other medication classes. Constructing a prescription network enriches the exploratory analysis of medication use at the population level.