### 2024-11-06 docs/index.qmd
Include the report directly to the index
This patch will avoid recursive file searching, which apparently introduces error when run with 'targets'.

---
knitr:
  opts_chunk:
    fig.width: 18
    fig.height: 12
---

```{r}

pkgs <- c("targets", "magrittr")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)
options(digits = 2, scipen = 999)

```

# Executive Summary {-}

:::: {.content-visible when-meta="in.report"}

## Analytical procedures

{{< include exec-summary/_source-data.qmd >}}

{{< include exec-summary/_preprocessing.qmd >}}

{{< include exec-summary/_measure-eigen.qmd >}}

{{< include exec-summary/_descriptive.qmd >}}

{{< include exec-summary/_seasonality.qmd >}}

{{< include exec-summary/_arima.qmd >}}

{{< include exec-summary/_spectral.qmd >}}

## Highlighted findings

{{< include exec-summary/_result-highlights.qmd >}}

::: {.content-visible when-profile="summary"}

\uselandscape

## Excerpt on findings

### Pair plot

```{r}

tar_read(plt_pair)[[17]]

```

### Dot plots on weekly time-series

```{r}

tar_read(plt_dot_week)[[1]]

```

### Differenced time-series on a weekly interval

```{r}

tar_read(plt_dot_diff_week)[[1]]

```

### Exploring seasonality within the number of medication claims

```{r}

tar_read(plt_period_month)[[1]]
tar_read(plt_acf_diff_month)[[1]]
tar_read(plt_pacf_diff_month)[[1]]

```

### Exploring ARIMA model

```{r}

tar_read(plt_arima_forecast_n_claim)[[17]]

```

### Decomposed time-series with SSA

```{r}

tar_read(plt_ssa_recon_n_claim)[[17]]

```

\normalpapersize

:::

::::

::: {.content-visible when-profile="descriptive"}
{{< include exec-summary/_descriptive.qmd >}}
:::

::: {.content-visible when-profile="arima"}
{{< include exec-summary/_arima.qmd >}}
:::

::: {.content-visible when-profile="seasonality"}
{{< include exec-summary/_seasonality.qmd >}}
:::

::: {.content-visible when-profile="spectral"}
{{< include exec-summary/_spectral.qmd >}}
:::

::: {.content-visible when-profile="models"}
{{< include exec-summary/_models.qmd >}}
:::

:::: {.content-visible when-meta="in.report"}

## Contents

- @sec-descriptive contains a descriptive explanation of the dataset
- @sec-seasonality reports exploration of seasonality on detrended data
- @sec-arima showcases results obtained from ARIMA models
- @sec-spectral outlines the decomposed time series using singular-spectrum analysis

## References

::: {#refs}
:::

::::