### 2024-11-26 docs/supplementary.qmd
Move the daily-variation figure to the supplementary #24

```{r}

pkgs <- c("targets", "magrittr")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)
options(digits = 2, scipen = 999)

```

```{r fig-period-day}
#| fig-width: 14
#| fig-height: 5
#| fig-cap: Daily cyclical pattern captured on a de-trended data

tar_read(plt_period_day)[[1]]

```