### 2024-12-10 docs/exec-summary/_measure-eigen.qmd
Revise the formal definition following the cited references

::: {.content-visible unless-meta="in.article"}
### Measuring relative importance
:::

As all medications are nodes in a claim network, we can measure the property of each medication using graph theory. One way to represent node property is by calculating node centrality, which can be measured as a number of incoming/outgoing edges (degree) or its connectedness to other nodes (importance). We measured importance using eigenvector centrality, which reflects medication co-occurrence with other important medications. Eigenvector centrality is a recursive measures of how well connected a certain node is to other nodes that is also well connected in the network [@Golbeck2013]. @eq-eigen-centrality outlines the calculation of eigenvector centrality; $c_i$ and $c_j$ are the centralities of node $\mathcal{v}_i$ and $\mathcal{v}_j$, respectively; $\lambda$ is the eigenvalue; $a_{ji}$ is an element on row $j$ and column $i$ from an adjacency matrix $A$, representing the connection between node $\mathcal{v}_j$ and $\mathcal{v}_i$.

$$
\displaystyle c_i = \frac{1}{\lambda} \sum_{j \neq i} a_{ji} \cdot c_j
$$ {#eq-eigen-centrality}