### 2024-12-10 docs/exec-summary/_preprocessing.qmd

::: {.content-visible unless-meta="in.article"}
### Data pre-processing
:::

Daily medication claims are represented as tabular data where each row corresponds to an individual patient’s prescriptions. The data was transformed into a square matrix summarizing co-prescriptions and single prescriptions to construct a graph-based representation. The rows and columns of the matrix represent groups of medications, while each matrix element indicates the frequency of a specific prescription pattern. Elements in the off-diagonal positions represent co-prescriptions between two medications, while diagonal elements represent single prescriptions. This matrix can be interpreted as a graph, where each medication is a node, and the edges represent co-prescription frequencies. The weight of the edge is derived from the number of daily claims. Originally, all defined daily doses (DDD) were assumed to equal 1, leading to straightforward edge weights based on raw counts. However, since not all medications have DDD values of 1, this assumption could lead to inaccurate representations. To address this, weights were adjusted using the density of a normal distribution. The expected value ($\mu$) was set to 1 to prioritize DDD = 1, and the standard deviation ($\sigma$) was set to $\frac{1}{3}$ to allow for a gradual reduction in weight as DDD deviates from 1. This adjustment ensures a more accurate weighting system that reflects real-world prescription patterns while preserving information about medications with non-standard DDD values.

$$\omega = \frac{1}{\frac{\omega_B}{3} \sqrt{2 \pi}} e^{-\frac{1}{2}\left(\frac{DDD - \omega_B}{\omega_B/3}\right)^2}$$ {#eq-weighted-DDD}

$$
\stackrel{\textrm{Unregularized}}{
  \begin{bmatrix}
  n_{1, 1} & \dots & n_{1, N} \\
  \vdots & \ddots & \vdots \\
  n_{N, 1} & \dots & n_{N, N} \\
  \end{bmatrix}
}
\quad \to \quad
\stackrel{\textrm{Regularized by } DDD}{
  \begin{bmatrix}
  \frac{1}{2} \cdot \left(\omega_1 + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_1 + \omega_N \right) \\
  \vdots & \ddots & \vdots \\
  \frac{1}{2} \cdot \left(\omega_N + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_N + \omega_N \right) \\
  \end{bmatrix}
}
$$ {#eq-regularize-mtx}