### 2025-03-12 docs/exec-summary/_descriptive.qmd
Revise the methods section

::: {.content-visible unless-meta="in.article"}
### Exploratory descriptive analysis
:::

The data was extracted as a daily time series and aggregated into weekly and monthly intervals for analysis. It includes four metrics: daily medication claims, daily patients, claim-to-patient ratio, and eigenvector centrality. For each metric, a univariate descriptive analysis was performed, reporting the mean, median, standard deviation, and interquartile range to summarize centrality and dispersion.

::: {.content-visible unless-meta="in.article"}
The data is split into pre-COVID-19 and during COVID-19 period, and mean difference was calculated using Welch's T-Test. For visual examination, pair plots were generated for each medication group. The correlations between metrics were calculated by considering pre-COVID-19, during COVID-19, and the overall association. Daily, weekly, and monthly time-series were also plotted and grouped by medications. Classical and seasonal-trend decomposition using LOESS (STL) were performed to extract decomposed trend for further visual evaluation. Each dataset was then detrended by performing a first-degree differencing, i.e. taking the differences between $x_t$ and $x_{t-1}$.
:::