### 2025-05-02 docs/exec-summary/_source-data.qmd

::: {.content-visible unless-meta="in.article"}
### Source of data
:::

All data in this study originated from IADB.nl, a dynamic pharmacy database containing prescription data from 1994 which still grow expansively. The database covers approximately 120 community pharmacies serving over one million patients and was accessed on the 20^th^of September, 2023. All patients are recorded in the database, irrespective of health care insurance, where the prescription rates, age, and gender are generalizable to represent the Netherlands [@Visser2013]. The database offers a longitudinal record, where each individual has a uniquely anonymized tracking identifier. All records contain fields on the date of medication dispensing, its quantity, dose regimen, number of days prescribed, prescribing physicians, and the Anatomical Therapeutic Chemical (ATC) code. Except for the over the counter drugs and medication dispensed during hospitalization, other medication records for each patient are complete, mainly due to high patient-pharmacy commitment in the Netherlands. This analysis includes daily medication dispensing from a static cohort of adults aged 18-65 years old and prescribed for anxiolytics or antidepressants at least once in the time span from 2018 to 2022.