### 2025-05-07 docs/article.qmd

---
title: >
  A temporal network analysis of drug co-prescription during antidepressant and anxiolytic dispensing in the Netherlands from 2018 to 2022

format:
  elsevier-pdf:
    keep-tex: true
    include-in-header:
      file: _header.tex
    journal:
      formatting: review
      cite-style: authoryear
    output-file: iadb-network-analysis.pdf
  docx:
    pandoc-options: ["--mathml"]
    output-file: iadb-network-analysis.docx

funding: "The author(s) received no specific funding for this work."
keywords:
  - Graph Theory
  - Prescription Registry
  - Co-prescription
  - Mental Health   
  - Public Health Monitoring

metadata-files:
  - _author.yml
  - _abstract-short.yml
  - _knitr-bib.yml

in:
  article: true
---

```{r}

pkgs <- c("targets", "magrittr")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)
options(digits = 2)

```

# Introduction

::: {.content-hidden}

Mental disorders are a common comorbidity affecting nearly two billion people worldwide with a prevalence that remains relatively stable overtime [@Kessler2009; @Steel2014; @Richter2019]. Having a chronic mental disorder is one factor which may lead to polypharmacy, *viz.* the use of multiple medications in treating complex clinical conditions [@Halli2019]. Polypharmacy is a prominent issue in modern medicine with a prevalence of 37% [@Delara2022], mainly attributed to the presence of multiple long-term comorbidity, i.e. multimorbidity, in the population. In resolving multimorbidity, the general edict of the evidence-based medicine is to follow an established clinical guideline to treat a particular clinical condition. However, these guidelines are tailored for a single disease without considering multimorbidity, resulting in the use of one or more medications per condition, leading to polypharmacy regimens for one patient with multiple conditions [@Molokhia2017]. Albeit clinical improvements offered by polypharmacy, risks of drug interaction and adverse effects also become more apparent, engendering a public health issue. Hitherto, no indicator of polypharmacy has been universally standardized [@Sirois2016], where the number of prescription is merely relied on as a crude metric [@Molokhia2017; @Delara2022].

:::

The inherently complex nature of health-related phenomena have fueled the growing implementation of network analysis, making it an important research framework in medicine and epidemiology [@Luke2007]. Past implementations of network analysis in public health mostly focused on transmission network, social network, and organizational network. Only recently, @Cavallo2012 demonstrated a novel framework to describe medication use on a population level, i.e. drug prescription network (DPN), a term which later is coined by @Bazzoni2015. Network analysis and graph theory offer an approach to model the relationship among building blocks of a complex system. From a complex system standpoint, polypharmacy is a prescription of multiple drug combinations which can be modelled as a network [@Miglio2021]. DPN can capture the complex relationships of a system that may vary over time, which is highly relevant to drugs prescribed for mental disorders.

In psychiatry, polypharmacy is defined as a patient with at least two concurrent psychiatric medications [@Shrivastava2013]. With a prevalence between 13-90%, psychiatric polypharmacy manifests in five categories, namely same-class, multi-class, adjunctive, augmentation, and total polypharmacy [@Shrivastava2013]. Same-class polypharmacy is the use of multiple medications from the same class. Multi-class polypharmacy is the use of multiple medications from different classes indicated for the same symptom cluster. Adjunctive polypharmacy is the use of additional medications to treat side effects due to other medications. Augmentation polypharmacy is the use of full-dose and sub-dose medications from a different class for the same symptom cluster. Finally, total polypharmacy is the total count of all medications used by a patient, both psychopharmaca and other classes alike [@Shrivastava2013]. For policymakers, differences in defining polypharmacy pose a challenge to design a monitoring system [@Sirois2016].

::: {.content-hidden}

Monitoring polypharmacy regimens requires a holistic approach since drug prescription highly depends on the physician's prescribing habits [@Shrivastava2013], patient-physician relationship, and pharmaceutical market, which all bear an interplay with the decision-making process [@Cavallo2012]. A holistic polypharmacy monitoring system needs to describe the importance of polypharmacy by incorporating both qualitative and quantitative measures [@Sirois2016]. Quantitatively attributing importance to each medication is challenging, accounting for polypharmacy as a complex and interlinked medication uses. Network analysis is a viable framework for analyzing polypharmacy. Constructing DPNs from drug prescription data has been previously investigated, where DPN-based approach represented the general trend in co-prescription [@Cavallo2012; @Bazzoni2015].

:::

A DPN is a graphical representation where medications correspond to nodes, and edges represent co-prescription relationships between them. Typically, simple undirected graphs are represented by a square symmetric adjacency matrix $A$, where each nonzero element $A_{ij}$ indicates that medications $i$ and $j$ were co-prescribed on a given day. A population-level DPN can be constructed by aggregating individual patient networks through matrix addition [@Cavallo2012]. Analyzing the structural characteristics of a DPN using both local (medication-level) and global (network-level) graph measures provides insights into co-prescription dynamics.

Node centrality measures, such as eigenvector centrality, quantify the importance of a medication in polypharmacy patterns [@Askar2021]. Addressing concerns about standardized polypharmacy indicators raised by @Sirois2016 and @Delara2022, we highlight eigenvector centrality as a key measure of a medication’s connection to highly prescribed drugs. A high eigenvector centrality suggests frequent co-prescription with commonly used medications, reflecting prescribing trends. Building on prior studies [@Cavallo2012; @Bazzoni2015], we hypothesize that eigenvector centrality serves as a meaningful quantitative indicator of polypharmacy. Therefore, this study aimed to assess the structural characteristics of temporal DPNs, focusing on the role of eigenvector centrality in identifying associated medications in adult patients prescribed for antidepressant or anxiolytic.

::: {.content-hidden}
> **"How do we implement a network analysis approach to quantitatively attribute polypharmacy?"**
:::

# Methods

## Source of data and study population

{{< include exec-summary/_source-data.qmd >}}

## Graph theory

Mathematically, a graph $\mathcal{G}$ is a pair $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, where $\mathcal{V}$ is the set of nodes and $\mathcal{E}$ is the set of edges. In a DPN, the node set $\mathcal{V} = \{\mathcal{v}_1, \mathcal{v}_2, \dots, \mathcal{v}_n\}$ represents a finite set of $n$ medications. Additionally, $\mathcal{E}$ is a finite set of edges, where each edge is a pair of nodes that corresponds to a co-prescription. In an undirected graph, the order of $i, j$ does not matter. Typically, undirected graphs do not contain self connections, $\{\mathcal{v}_i, \mathcal{v}_i\} \notin \mathcal{E}$ [@estrada2012structure].

A graph can be represented as an adjacency matrix $A$. The adjacency matrix $A$ has dimensions of $|\mathcal{V}| \times |\mathcal{V}|$, where $|\mathcal{V}|$ is the number of nodes in a graph. Each element of $A_{ij}$ is the edge, where in an unweighted graph, $A_{ij} = 1$ if a connection exists between $\mathcal{v}_i$ and $\mathcal{v}_j$, and $A_{ij} = 0$ otherwise. In a weighted graph, $A_{ij}$ can take any real value ($A_{ij} \in \mathbb{R}$) depending on the application [@estrada2012structure].

## Data pre-processing to build the data matrix

{{< include exec-summary/_preprocessing.qmd >}}

## Centrality measures in graph theory

Centrality measures formalize the identification of important nodes in a graph [@estrada2012structure]. Different node centrality measures quantify different structural properties of a node. Previous work on DPNs highlighted four centrality measures that could be used to assess the importance of a medication within a co-prescription network [@Miglio2021]. Degree centrality in a DPN describes the number of co-prescription with the medication of interest. High (low) degree centrality means the medication is often (seldom) co-prescribed. Betweenness centrality indicates the frequency of a medication connecting two other medications by the shortest possible path. High (low) betweenness centrality means the medication is (not) a "bridge" between different kind of medications. Closeness centrality is the average distance between one medication to all other medications in the DPN. High (low) closeness centrality means that the medication is (not) commonly co-prescribed. Eigenvector centrality reflects the number of co-prescription with medications that have vital role in the DPN. High (low) eigenvector centrality means that the medication is often (seldom) co-prescribed with other important medications.

The choice of centrality largely depends on the objective of network analysis. As a general guide, degree centrality is useful to identify popular medication and monitor drug overuse. Betweenness centrality is suitable for targeting for drug-interaction study and optimizing therapy plan. Closeness centrality indicates widely-used key medications and efficiency in treatment networks. Eigenvector centrality is helpful to identify influential medications and narrow down high-impact medications for drug monitoring. We defined high and low eigenvector centrality relative to the expected value $\frac{1}{n}$, based on the null model of uniform connectivity (see "Determining relative importance"). This study focused on eigenvector centrality to evaluate which medications have a significant influence on how antidepressants and anxiolytics are prescribed. @eq-eigen-centrality outlines the calculation of eigenvector centrality; $c_i$ and $c_j$ are the centralities of node $\mathcal{v}_i$ and $\mathcal{v}_j$, respectively; $\lambda$ is the eigenvalue; $A_{ji}$ is an element on row $j$ and column $i$ from an adjacency matrix $A$, representing the connection between node $\mathcal{v}_j$ and $\mathcal{v}_i$.

$$
\displaystyle c_i = \frac{1}{\lambda} \sum_{j \neq i} A_{ji} \cdot c_j
$$ {#eq-eigen-centrality}

## Data analysis

### Exploratory descriptive analysis

{{< include exec-summary/_descriptive.qmd >}}

### Exploring seasonality in the dataset

{{< include exec-summary/_seasonality.qmd >}}

### Summarizing co-prescription

This study explored same-class, multi-class, and total co-prescription. The summary statistics were based on prescription-day data aggregated at the population level over the year. Prescription-day is the number of medications dispensed on a given day. For example, if seven medications are prescribed for one day, it accounts for 7 prescription-days. The mean and standard deviation represent the average number of prescription-days per person for any co-prescription regimen.

### Decomposition with singular spectrum analysis

Classical and seasonal-trend decomposition techniques may not fully capture complex periodic patterns in time-series data. Singular spectrum analysis (SSA) is a non-parametric method that leverages Hankel matrix embedding for decomposition. In this study, SSA was applied to identify trends and seasonal components in weekly aggregated prescription data. To better isolate trend and seasonal components, sequential SSA was applied. First, a basic SSA model with the lag parameter $L$ = 52 was used to extract the dominant trend. The residuals were then analyzed with a second SSA model using $L = \frac{N}{2}$ to capture complex periodic patterns. The extracted trend, residuals, and oscillatory components were evaluated visually and statistically, with Mann-Kendall trend tests applied to assess long-term changes. A complete documentation of this approach and its rationale was included in the supplementary file section S1.1.

### Determining relative importance

Since the sum of all eigenvector centralities $c_i$ in equation [-@eq-eigen-centrality] equals 1, the expected eigenvector centrality for each node in a uniformly connected network is $\frac{1}{n}$, where $n$ is the total number of nodes. In such a network, each node has an equal probability of being connected to any other, and thus no node is more "influential" than another. This expected value value, $\frac{1}{n}$, serves as a baseline under a null model of uniform connectivity. Assuming eigenvector centrality scores approximately follow a normal distribution, we applied a one-sample Student's T-test with Bonferroni correction to assess how much each eigenvector centrality differs from the expected values. Nodes with $c_i$ greater than $\frac{1}{n}$ were categorized as having high eigenvector centrality, and those below the expected value were considered low.

# Results

## Description of the population

```{r}

tar_load(res_tbl_overview)

n_patient <- prettyNum(res_tbl_overview$total, big.mark = ",", scientific = FALSE)
hi_claim <- res_tbl_overview$overview |> dplyr::slice_max(mean_claim)
lo_claim <- res_tbl_overview$overview |> dplyr::slice_min(mean_claim)

hi_mean <- prettyNum(hi_claim$mean_claim, big.mark = ",")
lo_mean <- prettyNum(lo_claim$mean_claim, big.mark = ",")

```

IADB recorded `r n_patient` patients having at least one dispensing of antidepressants or anxiolytics within five years of data extraction. @tbl-overview-patient captures the demographical dynamics of the population from 2018 to 2022. The ratio of male to female only varied slightly, and the average age steadily increased over the year. These findings imply that the population demography stays relatively stable overtime without any indication of sudden changes.

```{r tbl-overview-patient}
#| tbl-cap: "Number of participating patients with at least one dispensing of antidepressants or anxiolytics from 2018 to 2022"

tar_read(res_tbl_patient) |>
  dplyr::mutate(
    "year"       = as.character(year),
    "Total"      = n_male + n_female,
    "male"       = sprintf("%s (%s)", prettyNum(n_male,   big.mark = ","), scales::percent(p_male)),
    "female"     = sprintf("%s (%s)", prettyNum(n_female, big.mark = ","), scales::percent(p_female)),
    "male_age"   = sprintf("%.2f (%.2f)", age_male, age_sd_male),
    "female_age" = sprintf("%.2f (%.2f)", age_female, age_sd_female)
  ) |>
  dplyr::select(year, male, female, Total, male_age, female_age) |>
  flextable::flextable() |>
  flextable::set_header_labels(
    year       = "Year",
    male       = "Male",
    female     = "Female",
    male_age   = "Male",
    female_age = "Female"
  ) |>
  flextable::add_header_row(
    values    = c("", "Number of Patients (%)", "Mean Age (SD)"),
    colwidths = c(1, 3, 2),
    top       = TRUE
  ) |>
  flextable::align(align = "center", part = "all") |>
  flextable::align(j = 1, align = "left", part = "all") |>
  #flextable::font(fontname = "Times", part = "all") |>
  flextable::fontsize(size = 7, part = "all") |>
  flextable::autofit(add_w = 0, add_h = 0) |>
  flextable::set_table_properties(opts_pdf = list("arraystretch" = 1, "tabcolsep" = 0))

```

## Cyclicality and seasonality of medication dispensing

Visual inspection of daily dispensing volumes revealed consistent weekly cycles. Dispensing peaked on `r hi_claim$day`, with an average of `r hi_mean` [SD: `r hi_claim$sd_claim`] and dropped to its lowest on `r lo_claim$day`, with an average of `r lo_mean` [SD: `r lo_claim$sd_claim`] dispensing. The variation between weekends and weekdays suggested a regular and repeating structure in the data. A large margin of difference between the highest and lowest average number of dispensing implied daily variation and cyclicality, where it peaks on Monday, starts declining afterwards, and is at the lowest during weekends, as shown in @tbl-overview-daily. The cyclical patterns are detailed in the supplementary material section S2.1. Seasonality was explored as dependencies at daily and weekly lags and further described in the supplementary material section S2.2.

```{r tbl-overview-daily}
#| tbl-cap: "Daily dispensing from 2018 to 2022 among patients with at least one dispensing of antidepressants or anxiolytics"

res_tbl_overview$overview |>
  dplyr::mutate(
    "mean_sd" = sprintf("%s (%.2f)", prettyNum(mean_claim, big.mark = ","), sd_claim),
    "median_IQR" = sprintf("%s (%.2f)", prettyNum(median_claim, big.mark = ","), IQR_claim),
    "Range" = sprintf("%s - %s", prettyNum(min_claim, big.mark = ","), prettyNum(max_claim, big.mark = ","))
  ) |>
  dplyr::select(day, mean_sd, median_IQR, Range) |>
  flextable::flextable() |>
  flextable::set_header_labels(
    day = "Day",
    mean_sd = "Mean (SD)",
    median_IQR = "Median (IQR)"
  ) |>
  flextable::add_header_row(
    values    = c("", "Number of Dispensing"),
    colwidths = c(1, 3),
    top       = TRUE
  ) |>
  flextable::align(align = "center", part = "all") |>
  flextable::align(j = 1, align = "left", part = "all") |>
  #flextable::font(fontname = "Times", part = "all") |>
  flextable::fontsize(size = 7, part = "all") |>
  flextable::autofit(add_w = 0, add_h = 0) |>
  flextable::set_table_properties(opts_pdf = list("arraystretch" = 1, "tabcolsep" = 0))

```

## Co-prescription in the population

@tbl-res-poly outlines the type of co-prescription in the population. Two statistics are reported in @tbl-res-poly: the sum and the mean with its standard deviation. The sum indicates the total prescription-days of co-prescription within a year. Thus, we can read the findings as follow:

> "In 2018, same-class co-prescription of antidepressants occurred for a total of 73,528 prescription-days across the population, with an average of 0.06 prescription-days per person. For every 100 people prescribed with antidepressants, there were approximately 6 days in total during which same-class co-prescription regimens occurred."

```{r tbl-res-poly}
#| tbl-cap: Summary statistics of different psychiatric co-prescription categories

tar_read(res_tbl_poly) |>
  dplyr::mutate(
    "year" = as.character(year),
    "mean_sd" = sprintf("%.2f [%.2f]", mean, sd),
    "median_iqr" = sprintf("%.2f (%.2f)", median, IQR)
  ) |>
  dplyr::select(type:medication, sum, mean_sd) |>
  tidyr::pivot_wider(names_from = type, values_from = c(sum, mean_sd)) |>
  dplyr::select(year, medication, sum_same, sum_multi, mean_sd_same, mean_sd_multi, dplyr::contains("total")) |>
  flextable::flextable() |>
  flextable::set_header_labels(
    year = "Year",
    medication = "Medication",
    sum_same = "Same-class",
    sum_multi = "Multi-class",
    mean_sd_same = "Same-class",
    mean_sd_multi = "Multi-class",
    sum_total = "N",
    mean_sd_total = "Mean [SD]"
  ) |>
  flextable::add_header_row(
    values    = c("", "Population Level", "Per Person (Mean [SD])", "Total Co-prescription"),
    colwidths = c(2, 2, 2, 2),
    top       = TRUE
  ) |>
  flextable::footnote(
    i = 1,
    j = c(3, 5),
    ref_symbols = c("1", "2"),
    part = "header",
    value = flextable::as_paragraph(
      c(
        "The total occurrence of co-prescription on the population level/perscription-day",
        "The expected occurrence of co-prescription per person/perscription-day",
        "Prescription day is defined as the number of dispensed medication per day",
        "Multi-class co-prescription is defined as the co-prescription of medications from different ATC classification"
      )
    )
  ) |>
  flextable::align(align = "center", part = "all") |>
  flextable::align(j = 1:2, align = "left", part = "all") |>
  flextable::fontsize(size = 7, part = "all") |>
  #flextable::font(fontname = "Times", part = "all") |>
  flextable::width(j = c(1:8), c(1, rep(1.8, 7)), unit = "cm") |>
  flextable::set_table_properties(opts_pdf = list("arraystretch" = 1, "tabcolsep" = 0))

```

## Time-series decomposition

For brevity, @fig-ssa only depicts the decomposition result of antidepressants dispensing over the years. SSA-based decomposition separates the trend from its harmonics, as shown in the lower half of the plot. The trend explains roughly 90% of the variability, and harmonics explain 10%. The first two harmonics, F1 and F2, captured most of the variability in a series compared to the rest of it. The trend, F1, and F2 were used to reconstruct the time series. The largest graph in @fig-ssa represents the reconstructed time-series (green line) and its original data points (gray line). Other decomposition results are reported in the supplementary file section S2.3.

```{r fig-ssa}
#| fig-width: 16
#| fig-height: 12
#| fig-pos: H
#| fig-cap: Time-series decomposition of antidepressants dispensing data using singular spectrum analysis-based approach

tar_read(plt_ssa_recon_n_claim)[[17]]

```

## Eigenvector centrality measures

```{r}

tar_load(res_recon_stat)

antidepressants <- res_recon_stat %>% subset(.$group == "Antidepressants")
anxiolytics     <- res_recon_stat %>% subset(.$group == "Anxiolytics")

respiratory <- res_recon_stat %>% subset(.$group == "Respiratory")
analgesics  <- res_recon_stat %>% subset(.$group == "Analgesics")

reportMedClaim <- function(med_obj) {
  res <- with(med_obj, prettyNum(n_claim, big.mark = ","))

  return(res)
}

reportMedEigen <- function(med_obj) {
  res <- with(med_obj, sprintf("%.2e [SD: %.2e]", eigen_mean, eigen_sd))

  return(res)
}

diffMed <- function(med_obj1, med_obj2, FUN) {
  med_rep1 <- FUN(med_obj1)
  med_rep2 <- FUN(med_obj2)

  res <- sprintf("%s vs %s", med_rep1, med_rep2)

  return(res)
}

```

Seven medication classes exhibited high centrality, as shown in @fig-hi-eigen and @tbl-desc. Notably, highly dispensed medications generally had high eigenvector centrality, consistent with the theoretical expectations from @eq-eigen-centrality. However, this relationship was not strictly proportional. For example, although antidepressants had a higher number of dispensing compared to medications for the respiratory system (`r diffMed(antidepressants, respiratory, reportMedClaim)`), they exhibited lower eigenvector centrality (`r diffMed(antidepressants, respiratory, reportMedEigen)`). A similar pattern was observed for anxiolytics, which had more dispensing than analgesics (`r diffMed(anxiolytics, analgesics, reportMedClaim)`) but lower eigenvector centrality (`r diffMed(anxiolytics, analgesics, reportMedEigen)`). This reflects that eigenvector centrality captures not just volume but also the influence of connection sin the network, highlighting that commonly dispensed drug within a population may not necessarily occupy the most central network positions.

```{r tbl-desc}
#| tbl-cap: Descriptive statistics of dispensing data from 2018 to 2022, ordered by Eigenvector centrality

res_recon_stat |>
  dplyr::arrange(-eigen_mean) |>
  dplyr::mutate(
    "mean_sd_eigen"     = sprintf("%.2e [%.2e]", eigen_mean, eigen_sd),
    "median_IQR_eigen"  = sprintf("%.2e [%.2e]", eigen_median, eigen_IQR),
    "mean_sd_dose"      = sprintf("%.2e [%.2e]", dose_mean, dose_sd),
    "median_IQR_dose"   = sprintf("%.2e [%.2e]", dose_median, dose_IQR),
    "mean_sd_weight"    = sprintf("%.2e [%.2e]", weight_mean, weight_sd),
    "median_IQR_weight" = sprintf("%.2e [%.2e]", weight_median, weight_IQR),
    "range_dose"        = sprintf("%.2e - %.2e", dose_min, dose_max),
    "range_weight"      = sprintf("%.2e - %.2e", weight_min, weight_max)
  ) |>
  subset(select = c(group, n_claim, mean_sd_eigen, mean_sd_dose, mean_sd_weight)) |>
  flextable::flextable() |>
  flextable::set_header_labels(
    group            = "",
    n_claim          = "Dispensing",
    mean_sd_eigen    = "Centrality",
    mean_sd_dose     = "DDD",
    mean_sd_weight   = "Weight"
  ) |>
  flextable::add_header_row(
    values    = c("", "Mean [SD]"),
    colwidths = c(2, 3),
    top       = TRUE
  ) |>
  flextable::align(align = "center", part = "all") |>
  flextable::align(j = 1, align = "left", part = "body") |>
  #flextable::font(fontname = "Times", part = "all") |>
  flextable::fontsize(size = 7, part = "all") |>
  flextable::autofit(add_w = 0, add_h = 0) |>
  flextable::set_table_properties(opts_pdf = list("arraystretch" = 1, "tabcolsep" = 0))

```

```{r fig-hi-eigen}
#| fig-cap: Clusters of medication with a significant co-prescription patterns on antidepressant and anxiolytic dispensing
#| fig-width: 9
#| fig-height: 10

tar_read(plt_hi_eigen_box)

```

# Discussion

This network analysis of dispensing data in individuals using antidepressants or anxiolytics revealed distinct co-prescription patterns. While same-class co-prescription was more common among antidepressant users, multi-class co-prescription predominated among those receiving anxiolytics. Notably, in both groups, multi-class co-prescription was at least ten times more prevalent than same-class co-prescription. Seven medication classes exhibited high eigenvector centrality, indicating their strong influence on co-prescription dynamics. These include medications for the alimentary tract and metabolism (ATC code `A`), blood and blood-forming organs (`B`), cardiovascular system (`C`), respiratory system (`R`), and analgesics (`N02`). These classes do not merely appear frequently but also serve as key connectors within prescribing networks, linking psychiatric treatments to chronic disease management. Their centrality suggests that they play a role in structuring multi-class co-prescription, either as standard co-prescriptions due to clinical guidelines or as necessary adjuncts in managing comorbidities or side effects, e.g., ATC `A` class.

The predomination of multi-class co-prescription among anxiolytics recipients is consistent with the recommendations in the Dutch Huisartsengenootschap [NHG; @NHG2019Anxiety]. Specifically, the NHG guideline for anxiety disorders advises short-term benzodiazepine use, limited to two to four weeks, as an adjunct to Selective Serotonin Reuptake Inhibitors (SSRIs) during treatment initiation. In contrast, the NHG guideline for depressive disorders recommends initiating pharmacotherapy with a single SSRI and discourages the combination of multiple antidepressants [@NHG2023Depression]. Therefore, the higher rate of multi-class co-prescription observed among anxiolytic users is consistent with evidence-based clinical practice, rather than an artifact of sampling or measurement methods.

Our findings also align with previous research showing that antidepressants and anxiolytics are frequently co-prescribed with other medications [@Shrivastava2013], and that multi-class regimens contribute most to psychopharmaca polypharmacy [@de2004polypharmacy]. Additionally, our study offers a more granular view of these patterns by characterizing their network properties. By distinguishing between same-class and multi-class co-prescriptions and evaluating medication centrality, we discover that certain drug classes function as hubs within the prescribing landscape. These characteristics likely reflect the underlying burden of multimorbidity in the population, where the bidirectional associations between depression/anxiety and chronic illnesses is well established [@qi2024longitudinal]. Conditions such as diabetes mellitus, thyroid disorders, and asthma have been linked to an increased risk of depression [@jang2024temporal]. The need for long-term pharmacological management further drives co-prescription patterns. Additionally, chronic illness-related anxiety can contribute to heightened prescribing of anxiolytics [@lebel2020health]. This underscores the interconnected nature of psychopharmaca prescribing patterns and highlights the complexity of medication management in these populations.

DPN complements traditional drug utilization analyses by capturing the structural properties of prescribing behaviors. Beyond evaluating individual drug utilization or basic pairwise co-prescriptions, DPN provides a structural perspective by mapping the interconnections between medications within a broader prescribing network [@Bazzoni2015]. The network approach is a particularly useful method for understanding complex prescribing dynamics, which allows for the identification of central medication classes that influence co-prescription patterns. The high centrality of certain medication classes suggests they serve as critical connectors in treatment regimens, providing the basis for further monitoring and evaluation.

By mapping the relationships between medications, DPN enables the detection of patterns that would be challenging to observe through standard statistical methods. For instance, the high centrality of certain medication classes suggests that they play a pivotal role in multi-class co-prescription, likely due to their involvement in managing both psychiatric and non-psychiatric conditions. As such, network analysis is a suitable approach for generating hypotheses for further causal or predictive studies, which has also been thoroughly discussed by @Askar2021.

We acknowledge several limitations in this study. First, data aggregation at the population level resulted in the loss of individual-level information, limiting interpretation to population trends. Second, while the adjustment of weights based on DDD introduces a novel element to the analysis, it is important to note that DDD values are not always equal to one, as defined by the WHO ATC system. Our study adopted a DDD-based weighting approach to refine network edges, aligning with the recommendation of @Cavallo2012. The adjusted weighting penalizes values deviating significantly from 1, reducing their impact on the network. Future research should explore alternative weighting techniques, such as patient-level dosage adjustments, to enhance network precision.

Despite the limitations, our study demonstrates the utility of DPN as a powerful, data-driven approach to analyzing medication dispensing data. By modelling intricate co-prescription patterns, DPN filters medication classes based on network properties, offering insights into prescribing behaviors. These insights can reveal important trends in medication use, which can be useful to pharmaceutical reviews and public health monitoring. By capturing the complexity of drug prescription relationships, DPN holds significant potential to improve decision-making in both clinical and administrative contexts.

Future DPN studies on individuals using antidepressants or anxiolytics should adopt a more fine-grained ATC classification to identify specific medications driving co-prescription patterns within the currently identified seven medication classes. A more granular approach could reveal whether certain drugs disproportionately contribute to multi-class co-prescription and whether these prescribing trends vary across patient demographics or healthcare settings. Such insights could refine our understanding of prescribing behaviors and inform targeted interventions for optimizing medication regimens.

::: {.content-hidden}
- Following @Cavallo2012, the weight of the adjacency matrix was the frequency of co-prescriptions or single prescriptions. We adjusted the weight to account for varying prescription dosages, alleviating previous limitation where the DDD was restricted to 1 [@Cavallo2012].
- Additionally, the predominance of multi-class polypharmacy in individuals using antidepressants or anxiolytics suggests the need for a more integrated approach to managing patients receiving psychotropic medications, ensuring that treatment decisions account for both psychiatric and non-psychiatric conditions.
:::

# Conclusion

This study demonstrates the utility of DPN in uncovering the patterns of co-prescriptions among antidepressant and anxiolytic recipients. We found that multi-class regimens were markedly more prevalent, particularly among anxiolytic users, consistent with clinical guidelines. Moreover, we identified seven highly central medication classes that act as hubs in the prescribing network, linking psychiatric medications to the management of chronic somatic conditions. These findings emphasize the role of multimorbidity in shaping co-prescription behaviors and highlight the importance of these seven classes in both psychiatric care and broader chronic disease management. Our study contributes a network-based perspective to drug utilization research, offering an alternative approach to identifying high-impact medication classes that warrant further monitoring and future investigation.

# References