### 2025-05-07 docs/supplementary.qmd

---
title: >
  Supplementary: A temporal network analysis of drug co-prescription around antidepressants and anxiolytics uses in the Netherlands from 2018 to 2022

format:
  pdf:
    keep-tex: true
    include-in-header:
      - file: _header.tex
      - text: |
          \renewcommand*{\thesection}{S\arabic{section}}
    output-file: iadb-network-analysis-supplementary.pdf
    number-sections: true
    number-depth: 3
    geometry:
      - inner=2cm
      - outer=3cm
      - top=2cm
      - bottom=2.5cm
  docx:
    pandoc-options: ["--mathml"]
    output-file: iadb-network-analysis-supplementary.docx

metadata-files:
  - _author.yml
  - _knitr-bib.yml

in:
  article: true
---

```{r}

pkgs <- c("targets", "magrittr")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)
options(digits = 2)

```

# Methods

## Decomposition with singular spectrum analysis

{{< include exec-summary/_spectral.qmd >}}

\uselandscape

# Results

## Cyclical patterns on de-trended data

```{r fig-period-day}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: Daily cyclical pattern captured on a de-trended data

tar_read(plt_period_day)[[1]]

```

\pagebreak

## Daily ACF and PACF plots

```{r fig-season-day}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: Daily cyclical pattern captured on a de-trended data

tar_read(plt_acf_day)[[4]]
tar_read(plt_pacf_day)[[4]]

```

\pagebreak

## Weekly ACF and PACF plots

```{r fig-season-week}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: Daily cyclical pattern captured on a de-trended data

tar_read(plt_acf_week)[[4]]
tar_read(plt_pacf_week)[[4]]

```

\normalpapersize

## Time-series decomposition of medication dispensing records

```{r}
#| fig-width: 16
#| fig-height: 9
#| fig-pos: H
#| fig-cap: Time-series decomposition using singular spectrum analysis-based approach

tar_read(plt_ssa_recon_n_claim)[[1]]
tar_read(plt_ssa_recon_n_claim)[[2]]
tar_read(plt_ssa_recon_n_claim)[[3]]
tar_read(plt_ssa_recon_n_claim)[[4]]
tar_read(plt_ssa_recon_n_claim)[[5]]
tar_read(plt_ssa_recon_n_claim)[[6]]
tar_read(plt_ssa_recon_n_claim)[[7]]
tar_read(plt_ssa_recon_n_claim)[[8]]
tar_read(plt_ssa_recon_n_claim)[[9]]
tar_read(plt_ssa_recon_n_claim)[[10]]
tar_read(plt_ssa_recon_n_claim)[[11]]
tar_read(plt_ssa_recon_n_claim)[[12]]
tar_read(plt_ssa_recon_n_claim)[[13]]
tar_read(plt_ssa_recon_n_claim)[[14]]
tar_read(plt_ssa_recon_n_claim)[[15]]
tar_read(plt_ssa_recon_n_claim)[[16]]
tar_read(plt_ssa_recon_n_claim)[[18]]
tar_read(plt_ssa_recon_n_claim)[[19]]
tar_read(plt_ssa_recon_n_claim)[[20]]
tar_read(plt_ssa_recon_n_claim)[[21]]
tar_read(plt_ssa_recon_n_claim)[[22]]
tar_read(plt_ssa_recon_n_claim)[[23]]
tar_read(plt_ssa_recon_n_claim)[[24]]

```

# References {-}