### 2025-07-10 docs/abstract/_abstract-short.qmd
Finalize the first draft

**Background:**
Drug prescription networks (DPNs) model the temporal dynamics of medication co-prescription within a population.  Understanding these networks can provide insights into polypharmacy and prescribing behaviors.

**Objectives:**
This study assesses the structural characteristics of temporal DPNs derived from daily co-prescriptions of antidepressants, anxiolytics, and other therapeutic drug classes. By analyzing these networks using eigenvector centrality, we identify influential medications and prescribing patterns.

**Methods:**
We utilized the IADB.nl database, including prescriptions from 128 Dutch pharmacies (2018–2022). A cohort of patients prescribed antidepressants/anxiolytics was extracted. Medications were classified using the Anatomical Therapeutic Chemical (ATC) system into 24 therapeutic classes. Time-varying DPNs were constructed as undirected graphs using symmetric daily dose-adjusted co-prescriptions. Eigenvector centrality ($c_e$) quantified relative nodal importance.  Weekly-aggregated data included prescriptions ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality. Singular spectrum analysis decomposed trends.

**Results:**
Antidepressants ($c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics ($c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) had high eigenvector centrality, demonstrating frequent co-prescription. Other ATC groups with high centrality included those for the alimentary tract and metabolism (`A01-A16`), blood and blood-forming organs (`B01-B06`), cardiovascular system (`C01-C10`), respiratory system (`R01-R07`), and analgesics (`N02`).

**Discussion:**
DPNs revealed key co-prescription patterns. High-centrality medications highlight potential targets for drug monitoring, such as identifying co-prescription trends that may warrant evaluation for safety, appropriateness, or policy oversight. This approach aids in identifying influential medications and refining prescribing oversight.