### 2025-07-26 docs/exec-summary/_preprocessing.qmd
Finalize the IJE draft

::: {.content-visible unless-meta="in.article"}
### Data pre-processing
:::

Assume an observational study where $P$ individuals indexed by $p$, $p=1,\dots,P$ are followed for each day $t$, $t=1, \dots, T$ for a total period of $T$ days. To derive the population-level DPN on each day $t$ with adjacency matrix $A_t$, the individual-level daily prescription matrices $A^{p}\left(t\right)$ were aggregated by computing their element-wise sum, that is,

$$
\displaystyle \sum_{p=1}^P A^{p}\left(t\right), \ \text{where} \ A\left(t\right) \in \ \mathbb{Z}_+^{n\times n}.
$$ {#eq-DPN}

The matrix $A\left(t\right)$ captures the population frequency of co-prescriptions (off-diagonal) and single prescriptions (diagonal) in the data at

the specified time $t$. Each element $A_{ij}\left(t\right)$ in the matrix was interpreted as the number of patients being prescribed medication classes $i$ and $j$, with $i,j\in \mathcal{V}$, on the same day $t$.

The population-level DPN is then a weighted undirected graph $\mathcal{G}\left(t\right) = \left(\mathcal{V}, \mathcal{E}\left(t\right)\right)$ observed over $T$ discrete time points $t=1,\dots,T$, where:
- $\mathcal{V}$ is the fixed set of $n$ nodes,
- $\mathcal{E}_t \subseteq \{ \{i, j\} : i, j \in \mathcal{V}, i \neq j \}$ is the edge set at time $t$.

The population-level DPN at time $t$ is fully described by its symmetric weighted adjacency matrix $A\left(t\right) \in \mathbb{Z}_+^{n \times n}$, which is defined as:

$$
{A}_{ij}\left(t\right) =
\begin{cases}
w_{t,ij} > 0 & \text{if } \{i, j\} \in \mathcal{E}\left(t\right), \\
0 & \text{otherwise},
\end{cases}
\quad \text{with } {A}\left(t\right) = {{A}\left(t\right)}^\top.
$$

In the original work of DPN, all defined daily dose (DDD) were assumed equal to 1, allowing for applying raw counts as the edge weights. However, since not all medications have DDD = 1, this assumption could lead to inaccurate representations. To enhance the accuracy of edge weights, we introduced a weighting function to adjust the contribution of each co-prescription. Specifically, each co-prescribed medication class was assigned a weight $\omega_j$ based on its DDD value using a Gaussian kernel centered at the baseline weight $\omega_B = 1$, i.e., the expected DDD, with standard deviation $\sigma = \frac{1}{3}$ to allow for a gradual reduction in weight as DDD deviates from 1.

$$\omega_j = \frac{1}{\sigma \sqrt{2 \pi}} e^{-\frac{\left(DDD_j - 1\right)}{2\sigma^2}^2}$$ {#eq-weighted-DDD}

The edge weight between medication classes $i$ and $j$ was computed as the average of their individual weights:

$$
\stackrel{\textrm{Unregularized}}{
  \begin{bmatrix}
  n_{1, 1} & \dots & n_{1, N} \\
  \vdots & \ddots & \vdots \\
  n_{N, 1} & \dots & n_{N, N} \\
  \end{bmatrix}
}
\quad \to \quad
\stackrel{\textrm{Regularized by } DDD}{
  \begin{bmatrix}
  \frac{1}{2} \cdot \left(\omega_1 + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_1 + \omega_N \right) \\
  \vdots & \ddots & \vdots \\
  \frac{1}{2} \cdot \left(\omega_N + \omega_1 \right) & \dots & \frac{1}{2} \cdot \left(\omega_N + \omega_N \right) \\
  \end{bmatrix}
}
$$ {#eq-regularize-mtx}

This weighting approach scaled the edge weight that deviate substantially from the standard $\omega_B = 1$, improving comparability across medication classes. To remain consistent with standard network representations and avoid confounding self-loops, the diagonal elements were omitted.